---
title: "Great Reed Warbler alarm call attracts brood parasites"
author: "Attila Marton, Attila Fülöp, Csaba Moskát, Miklós Bán"
date: '2018 június 7 '
output: html_document
---

Files:

DFx2b.csv contains  (1) great reed warbler (Acrocephalus arundinaceus, GRW from hereafter) nest checking data from 2013, 2014 and 2018, 
                    (2) great reed warbler nest characteristics from 2018
                    (3) male and female great reed warbler alarm call intensity data to observer and to dummy plastic cuckoo from 2018
                    (4) male and female common cuckoo (Cuculus canorus CC from hereafter) reactions to GRW alarm call from 2018. All experiment were performed near Apaj village, central Hungary.
                    
- Columns:
-- nestid: a concatenated id showing the number of given nest in given year and the year
-- id: the id number of given nest within given year
-- year: the year of the study
-- date: format yyyy.mm.dd
-- site: name of channel (Arapaszto, Apaj, DVCS, Kunszentmiklos)
-- grw.eggs: number of great reed warbler eggs at the time of visit
-- cc.eggs: number of common cuckoo eggs at the time of visit
-- total.eggs: grw.eggs + cc.eggs
-- pull: 1 if chicks hatched at the time of visit
-- abandoned: 1 if nest was deemed abandoned at the time of visit
-- laying.date: the presumed laying date of the first egg
-- observer: initials of observers
-- exp: 0 = nest checking, 1 = small clutch, 2 = full clutch, 3 = drone flight
-- boots.grwm: male GRW response to man in boots (0 = none, 1 = weak alarm call, 2 = strong alarm call)
-- boots.grwf: female GRW response to man in boots (0 = none, 1 = weak alarm call, 2 = strong alarm call)
-- playback.grwm: male GRW response to plastic cuckoo dummy during GRW alarm call playback (0 = none, 1 = weak alarm call, 2 = strong alarm call, 3 = contact fight)
-- playback.grwf: female GRW response to plastic cuckoo dummy during GRW alarm call playback (0 = none, 1 = weak alarm call, 2 = strong alarm call, 3 = contact fight)
-- before.ccm: male CC presence near the nest before GRW alarm call playback or natural GRW alarm call (0 = none, 1 = calling, 2 = flying closer, 3 = flying closer and vocalizing intensively)
-- before.ccf: female CC presence near the nest before GRW alarm call playback or natural GRW alarm call (0 = none, 1 = calling, 2 = flying closer, 3 = flying closer and vocalizing intensively)
-- during.ccm: male CC reaction near the nest to GRW alarm call playback or natural GRW alarm call (0 = none, 1 = calling, 2 = flying closer, 3 = flying closer and vocalizing intensively)
-- during.ccf: female CC reaction near the nest to GRW alarm call playback or natural GRW alarm call (0 = none, 1 = calling, 2 = flying closer, 3 = flying closer and vocalizing intensively)
-- after.ccm: male CC reaction near the nest after GRW alarm call playback or natural GRW alarm call (0 = none, 1 = calling, 2 = flying closer, 3 = flying closer and vocalizing intensively)
-- after.ccf: female CC reaction near the nest after GRW alarm call playback or natural GRW alarm call (0 = none, 1 = calling, 2 = flying closer, 3 = flying closer and vocalizing intensively)
-- pamela: 1 if dummy plastic cuckoo was used
-- alarm.file: the name of the alarm file used
-- record: if GRW reaction was recorded or not
-- obs: any observation related to the nest visit
-- nest.height: the height of the nest in cm
-- nest.width: the width of the nest at the top in cm
-- dist.water: distance of nest from water in cm
-- dist.bank: distance of nest from the bank in cm
-- above.water: distance between water and the top of the nest in cm
-- above.nest: distance from the top of the nest and the top of the vegetation in cm
-- visibility: direct nest view (DNV), indirect nest view (INV), no nest view (NNW) from the nest to the nearest perching tree
-- perch.dist: distance in m to the nearest perching tree

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library setup}
library(chron)
library(doBy)
library(rptR)
library(beepr)
library(Matrix)
library(lme4)
#library(nlme)
#library(ordinal)
#library(car)
#library(rgdal)
#library(geosphere)
```

```{r dataset and variable setup, echo=TRUE}

# setting up work directory

setwd("C:/Users/Marton Attila/OneDrive/PhD/GRW-alarm-call")

df <- read.csv("DFx2b.csv", header = T, sep = ";")
summary(df)

# converting nestid, year, site and observer to factor

df$nestid <- as.factor(df$nestid)
df$year <- as.factor(df$year)
df$site <- as.factor(df$site)
df$observer <- as.factor(df$observer)

# converting date to julian date in column j.date using pachage chron

df$date <- as.Date(df$date, format = '%Y.%m.%d')

df$j.date[df$year == "2013"] <- julian(df$date[df$year == "2013"], origin = as.Date("2013-01-01"), units = "days")

df$j.date[df$year == "2014"] <- julian(df$date[df$year == "2014"], origin = as.Date("2014-01-01"), units = "days")

df$j.date[df$year == "2018"] <- julian(df$date[df$year == "2018"], origin = as.Date("2018-01-01"), units = "days")

# converting laying date to julian date in column j.laying.date

df$laying.date <- as.Date(df$laying.date, format = "%Y.%m.%d")

df$j.laying.date[df$year == "2013"] <- julian(df$laying.date[df$year == "2013"], origin = as.Date("2013-01-01"))

df$j.laying.date[df$year == "2014"] <- julian(df$laying.date[df$year == "2014"], origin = as.Date("2014-01-01"))

df$j.laying.date[df$year == "2018"] <- julian(df$laying.date[df$year == "2018"], origin = as.Date("2018-01-01"))

# adding column nestday by subtracting j.date from j.laying.date

df$nestday <- df$j.date - df$j.laying.date

# converting time to minutes from midnight in column mminutes

df$time <- as.character(df$time)
df$time2 <- paste(df$time, ":00", sep = "")
df$time2[df$time2 == "NA:00"] <- NA
df$mminutes <- 60*24*as.numeric(times(df$time2))

# adding column parasite 

df$parasite[df$cc.eggs == "0"] <- 0
df$parasite[df$cc.eggs > "0"] <- 1
df$parasite <- as.factor(df$parasite)

# we had some concernes regarding the correct identification of class 1 alarm calls, so we made some histograms with the observers. It turns out GN probably cannot reliably identify class 1 alarm calls, which is not surprising, due to the ambiguity of the class 1. We decided to drop class 1 alarm calls from all further analyses. Thus, we have a new binomial variable: "0" is no alarm, "1" is (strong) alarm call.

par(mfrow=c(1,2))
hist(df$boots.grwm, main = "Hist of boots.grwm by all obs")
hist(df$boots.grwf, main = "Hist of boots.grwf by all obs")

par(mfrow=c(1,2))
hist(df$boots.grwm[df$observer == "ZSI"], main = "Hist of boots.grwm by ZSI")
hist(df$boots.grwf[df$observer == "ZSI"], main = "Hist of boots.grwf by ZSI")

par(mfrow=c(1,2))
hist(df$boots.grwm[df$observer == "BM"], main = "Hist of boots.grwm by BM")
hist(df$boots.grwf[df$observer == "BM"], main = "Hist of boots.grwf by BM")

par(mfrow=c(1,2))
hist(df$boots.grwm[df$observer == "BM, MA"], main = "Hist of boots.grwm by BM, MA")
hist(df$boots.grwf[df$observer == "BM, MA"], main = "Hist of boots.grwf by BM, MA")

par(mfrow=c(1,2))
hist(df$boots.grwm[df$observer == "MA"], main = "Hist of boots.grwm by MA")
hist(df$boots.grwf[df$observer == "MA"], main = "Hist of boots.grwf by MA")

par(mfrow = c(1,2))
hist(df$boots.grwm[df$observer == "GN"], main = "Hist of boots.grwm by GN")
hist(df$boots.grwf[df$observer == "GN"], main = "Hist of boots.grwf by GN")

dev.off()

# removing class 1 and converting class 2 to class 1
df$boots.grwf[df$boots.grwf == "0"] <- 0
df$boots.grwf[df$boots.grwf == "1"] <- NA
df$boots.grwf[df$boots.grwf == "2"] <- 1

df$boots.grwm[df$boots.grwm == "0"] <- 0
df$boots.grwm[df$boots.grwm == "1"] <- NA
df$boots.grwm[df$boots.grwm == "2"] <- 1

par(mfrow = c(1,2))
hist(df$boots.grwf, main = "Binomial boots.grwf")
hist(df$boots.grwm, main = "Binomial boots.grwm")
dev.off()

df$pair[df$boots.grwm == "0" & df$boots.grwf == "0"] <- 0
df$pair[df$boots.grwm == "1" & df$boots.grwf == "0"] <- 1
df$pair[df$boots.grwm == "0" & df$boots.grwf == "1"] <- 2
df$pair[df$boots.grwm == "1" & df$boots.grwf == "1"] <- 3
df$pair <- as.factor(df$pair)

# adding column bpair
df$bpair[df$pair == "0"] <- 0
df$bpair[df$pair != "0"] <- 1

# merging during.ccm with after.ccm and during.ccf with after.ccf

df$response.ccm <- NA

for (i in 1:length(df$response.ccm)) {
  
    if (!is.na(df$during.ccm[i])) {
        if (df$during.ccm[i] > df$after.ccm[i]) {
            df$response.ccm[i] <- df$during.ccm[i]
        } else {
            df$response.ccm[i] <- df$after.ccm[i]
        }
    }

}

df$response.ccf <- NA

for (i in 1:length(df$response.ccf)) {
  
    if (!is.na(df$during.ccf[i])) {
        if (df$during.ccf[i] > df$after.ccf[i]) {
            df$response.ccf[i] <- df$during.ccf[i]
        } else {
            df$response.ccf[i] <- df$after.ccf[i]
        }
    }

}


```

```{r subseting and removing NAs}

# creating a full df
df.full <- df
df <- df.full

# removing NAs and subseting to nest visitis between 6 AM and 12 PM
#df <- df[!is.na(df$parasite),]
#df <- df[!is.na(df$nestday),]
#df <- df[!is.na(df$mminutes),]
#df <- df[!is.na(df$site),]
#df <- df[!is.na(df$boots.grwf),]
#df <- df[!is.na(df$boots.grwm),]
#df <- df[!is.na(df$pair),]

df <- subset(df, mminutes >= 360 & mminutes <= 720 | mminutes >= 1000 )

# Time conversions:
# 360 = 06:00
# 720 = 12:00
# 660 = 11:00
# 1000 = 16:40
# 1020 = 17:00
# 1080 = 18:00
# 1140 = 19:00

#df <- df[df$observer != "ZA",]
#df <- df[df$year != "2014",]

#df <- subset(df, nestday > 0) # removing negative nestday values
df <- subset(df, pull = 1 ) # removing nests with pull
df <- subset(df, abandoned = 1) # removing abandoned nests

```

```{r exploratory analyses on df}

head(df)
summary(df)
str(df)

```

```{r repeatability of alarm call}

# repeatability with one grouping level

df$multiple <- 1
n <- summaryBy(multiple ~ nestid, data = df, FUN = c(sum))
m <- n$nestid[n$multiple.sum > 1]

df.rep <- df[df$nestid %in% m,]
df.rep <- df[!is.na(df$boots.grwf),]

boots.grwf.rpt <- rptBinary(boots.grwf ~ (1|nestid), grname=c("nestid"), data=df.rep)
beep(4)
boots.grwf.rpt

boots.grwm.rpt <- rptBinary(boots.grwm ~ (1|nestid), grname=c("nestid"), data=df.rep)
beep(4)
boots.grwm.rpt

pair.rpt <- rptBinary(bpair ~ (1|nestid), grname=c("nestid"), data=df.rep)
beep(4)
pair.rpt

# number of df.rep nests, mean and sd

sumvisit <- summaryBy(multiple ~ nestid, data = df.rep, FUN = c(sum))
length(unique(df.rep$nestid))
mean(sumvisit$multiple.sum)
sd(sumvisit$multiple.sum)

# percentage of nests with at least two visits

(length(unique(df.rep$nestid))*100)/length(unique(df$nestid))

```

```{r cor.test between boots.grw and playback.grw}

df.cor <- df[!is.na(df$playback.grwf),]

cor.test(df.cor$boots.grwf, df.cor$playback.grwf, method = "pearson")

```


```{r GRW alarm call probability to observers correlates to clutch size}

m1.1 <- glmer(boots.grwf ~ total.eggs
               + scale(mminutes)
               + scale(nestday)
               + scale(j.date)
               + scale(laying.date)
               + parasite
               + (1|nestid)
               + (1|observer)
               + (1|site)
               ,
               data=df, family = "binomial",
               control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)),
               na.action = na.omit)
summary(m1.1)
drop1(m1.1, test = "Chisq")

m1.2 <- update(m1.1, .~. - parasite)   
summary(m1.2)
drop1(m1.2, test = "Chisq")

m1.3 <- update(m1.2, .~. - scale(nestday))
summary(m1.3)
drop1(m1.3, test = "Chisq")

m1.4 <- update(m1.3, .~. - scale(mminutes))
summary(m1.4)
drop1(m1.4, test = "Chisq")

m1.5 <- update(m1.4, .~. - scale(laying.date))
summary(m1.5)
drop1(m1.5, test = "Chisq")

m1.6 <- update(m1.5, .~. - scale(j.date))
summary(m1.6)

```

```{r}

m2.1 <- glmer(boots.grwm ~ total.eggs
              + scale(mminutes)
              + scale(nestday)
              + scale(j.date)
              + scale(laying.date)
              + parasite
              + (1|nestid)
              + (1|observer)
              + (1|site),
              data = df, family = "binomial",
              control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)),
              na.action = na.omit              )
summary(m2.1)
drop1(m2.1, test = "Chisq")

m2.2 <- update(m2.1, .~. - scale(mminutes))
summary(m2.2)
drop1(m2.2, test = "Chisq")

m2.3 <- update(m2.2, .~. - scale(nestday))
summary(m2.3)
drop1(m2.3, test = "Chisq")

m2.4 <- update(m2.3, .~. - parasite)
summary(m2.4)
drop1(m2.4, test = "Chisq")

m2.5 <- update(m2.4, .~. - scale(laying.date))
summary(m2.5)

m2.6 <- update(m2.5, .~. - scale(j.date))
summary(m2.6)

```

```{r}

m3.1 <- glmer(parasite ~ playback.grwf
              + playback.grwm
              + total.eggs
              + scale(mminutes)
              + scale(nestday)
              + scale(j.date)
              + scale(laying.date)
              + parasite
              + (1|nestid)
              + (1|observer)
              + (1|site),
              data = df, family = "binomial",
              control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)),
              na.action = na.omit)
summary(m3.1)

m3.2 <- update(m3.1, .~. - scale(j.date) - scale(mminutes) - scale(laying.date)
               - boots.grwf - boots.grwm)
summary(m3.2)

```














